HPC On-boarding

### 2016-01-20 hpc_101/paper.tex
addition of globus walk through, and partial table of partition details

\author{Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.} % Your name
\institute[University of {\hawaii} -- ITS--CI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Information Technology Services \\
Department of {\ci} \\
University of {\hawaii} \\ % Your institution for the title page
\medskip
\textbf{\url{https://www.hawaii.edu/its/ci/}}\\
\textbf{\textit{uh-hpc-help@lists.hawaii.edu}} % Your email address
}
\date{\today} % Date, can be changed to a custom date

@@

\subsection{{\craycs}}
\begin{frame}
\frametitle{{\craycs} -- History}
\begin{itemize}
		\item Brought online Fall 2014 -- Spring 2015
		\item Initial investment of 1.8 Million by the University of {\hawaii} (UH)
		\item As of October 2015, more than 200 users have been granted access to the cluster
	\end{itemize}
\end{frame}

\subsubsection{Hardware}
\begin{frame}
	\frametitle{{\craycs}0 -- Compute Nodes}
	\begin{itemize}
		\item 3,800 total cores -- \textbf{Spring upgrade} $\rightarrow$ 5,400$+$
		\item 178 standard nodes -- \textbf{Spring upgrade} $\rightarrow$ 270
			\begin{itemize}
				\item Two 10 core {\intel} processors (\emph{20 cores total})
				\item Diskless -- Some RAM is used for the Operating System
				\item $\approx128$GB of useable RAM
			\end{itemize}
		\item 6 large memory nodes
			\begin{itemize}
				\item Four 10 core {\intel} processors (\emph{40 cores total})
				\item Diskless -- Some RAM is used for the Operating System
				\item $\approx1$TB of useable RAM
			\end{itemize}			
		\item CentOS Linux
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{{\craycs} -- Storage $\rightarrow$ {\lustre}}
	\begin{itemize}
		\item {\lustre} is a high performance parallel filesystem
		\item The {\craycs} has $\approx582$TB of storage space
		\item Shared between all compute nodes and login nodes
		\item Primarily used as scratch space for jobs (Input and Output)
		\item User do not have a usage quota (soft or hard)
		\item Certain directories are subject to a 90 day purge policy
		\item \textbf{Data is not backed up!  Users are responsible for their own data}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{{\craycs} -- Storage $\rightarrow$ ValueStorage}
	\begin{itemize}
		\item $500$TB of scale out storage
		\item Currently only available for purchase by cluster users and only accessible via the login nodes
		\item ValueStorage will eventually be available for purchase by everyone
		\item ValueStorage owners will eventually be able to mount as a network drive (CIFS) on laptop, workstations, servers
		\item Purchased in 0.5TB increments
	\end{itemize}
	\btVFill

@@

\begin{frame}
	\frametitle{{\craycs} -- Network}
	\begin{itemize}
		\item 40Gb Infiniband inter-connects (QDR)
		\begin{itemize}
			\item High speed inter-connect between~\\compute nodes, {\lustre} storage and Login nodes
			\item Utilizes the \emph{fat tree network topology}
			\item[] \includegraphics[width=0.20\textwidth]{images/Fat_tree_network} \\[-1ex] {\fontsize{3}{4} \selectfont Source: \url{https://en.wikipedia.org/wiki/Fat_tree} } 		
		\end{itemize}
		\item 10Gb login node internet connectivity
		\begin{itemize}
			\item Speed test from UH to CERN clocked transfer speeds up to 2$+$ Gb/s
		\end{itemize}	
	\end{itemize}
\end{frame}

@@

\section{Sustainability}
\subsection{Community Resources}
\begin{frame}
	\frametitle{Community Resources}
	\begin{itemize}
		\item The initial investment in the cluster provides resources for all Faculty, Staff, and Students affiliate with the University of {\hawaii}
		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, lm.q, sb.q, kill.q, htc.q}
		\item For some users, the publicly available resources may not be enough~\ldots
	\end{itemize}
\end{frame}

@@

\subsubsection{Service Units}
\begin{frame}	
	\frametitle{Service Units}
	\begin{itemize}
	 \item In some cases, users will not require owning a node, but still need priority access
	 \item An alternative to purchasing a node, is to purchase service units (\emph{SU})
	 \item SUs come in two variates:
		 \begin{itemize}
			 \item[--] Standard node units 
				\begin{itemize}
					\item 20 core hours, with access to 128GB of ram on a standard~node
				\end{itemize}
			 \item[--] Large memory node units
				\begin{itemize}
					\item 40 core hours, with access to 1TB of ram on a large~memory~node
				\end{itemize}
		 \end{itemize}
	 \item A minimum order totaling \$500 is required
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Globus}\footnotesize
If you already use Globus, the endpoints for the {\craycs} are:
\begin{itemize}
\item hawaii\#UHHPC1
\item hawaii\#UHHPC1
\end{itemize}
~\\ 
For those that have not used Globus, and wish to try it out, you can learn how to use Globus by visiting the {\ci} website for more information:~\\
{\footnotesize \href{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}}
\end{frame}

@@

\subsection{Managing User Jobs}
\begin{frame}
\frametitle{Managing User Jobs}
User jobs all come in different shapes and sizes:  
\begin{itemize}\footnotesize
	\item Require multiple nodes working in concert towards a common goal (MPI)
	\item Require a single node, in which they use multiple threads~\\work together (OpenMP, pthreads)
	\item Require a lot of cores to process a lot of data in an identical manner,~\\yet none of the inputs have dependencies on another (HTC)
\end{itemize}
\bigskip
The {\craycs} is capable of handling many different types of jobs, but with so many users in a multi-user environment, how do we turn this chaos into organized chaos?
\btVFill
\begin{center}This looks like a job for a \textbf{\emph{job scheduler}}!\end{center}
\end{frame}

\begin{frame}
	\frametitle{What is a Job Scheduler?}
	\begin{definition}{\footnotesize
		A job scheduler is a computer application for controlling unattended background program execution (commonly called batch processing)
		}
	\end{definition}
	Basic features expected of a job scheduler include:
	\begin{itemize}\footnotesize
    \item Interfaces which help to define workflows and/or job dependencies
    \item Automatic submission of executions
    \item Interfaces to monitor the executions
    \item Priorities and/or queues to control the execution order of unrelated jobs
	\end{itemize}
	For the {\craycs}, we utilize the \textbf{S}imple \textbf{L}inux \textbf{U}tility \textbf{R}esource \textbf{M}anager~\\ or simply known as the \emph{SLURM scheduler}
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Job_scheduler}}
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Slurm_Workload_Manager}}
	\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/slurm.html}}
\end{frame}

@@

\subsection{SLURM Partitions}
\begin{frame}
\footnotesize
\frametitle{Partitions}
\begin{block}{What is a partition?}\footnotesize
A partition can be thought of as a group of nodes/resources divided into possibly overlapping sets.  Each partition can be considered as a job queue, each of which has an assortment of constraints such as job size limit, job time limit, users permitted to use it, etc. Priority-ordered jobs are allocated nodes within a partition until the resources (nodes, processors, memory, etc.) within that partition are exhausted.\footnotemark
\end{block}
\begin{itemize}
	\item The {\craycs} currently has six public partitions:~\\\textbf{lm.q}, \textbf{community.q}, \textbf{exclusive.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
	\item More details on partition constraints can be viewed by visiting the \href{http://www.hawaii.edu/its/ci/hpc-resources/slurm-partitions/}{{\ci} website} or by using the following command
	\begin{itemize}\footnotesize
		\item[] \texttt{$[$login \ctilde$]$\$ scontrol show partition $<$partition name$>$}
		\item[] Example: \texttt{$[$login \ctilde$]$\$ scontrol show partition community.q}
	\end{itemize}

\end{itemize}
\footnotetext[2]{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
\end{frame}

@@

>>>>

\author{Sean Cleveland Ph.D, Ron Merrill Ph.D,\\David Schanzenbach M.S.} % Your name
\institute[University of {\hawaii} -- ITS--CI] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Information Technology Services \\
{\ci} \\
University of {\hawaii} \\ % Your institution for the title page
\medskip
\textbf{\url{https://www.hawaii.edu/its/ci/}}\\
\textbf{\textit{uh-hpc-help@lists.hawaii.edu}} % Your email address
}
\date{\today} % Date, can be changed to a custom date

@@

\subsection{{\craycs}}
\begin{frame}
\frametitle{{\craycs} -- History}
\begin{itemize}
 \item Initial investment of 1.8 Million by the University of {\hawaii} (UH)
 \item Deilvered in October of 2014
 \item Accepted in December of 2014 
 \item Early adopter testing started in January 2015
 \item Opened for general use in April 2015
 \item As of October 2015, more than 200 users have been granted access to the cluster
\end{itemize}
\end{frame}

\subsubsection{Hardware}
\begin{frame}
	\frametitle{{\craycs}0 -- Compute Nodes}
	\begin{itemize}
		\item 3,800 total cores -- \textbf{Spring upgrade} $\rightarrow$ 5,400$+$
		\item 178 standard nodes -- \textbf{Spring upgrade} $\rightarrow$ 270
			\begin{itemize}
				\item Two 10 core {\intel} processors (\emph{20 cores total})
				\item Diskless -- Some RAM is used for the Operating System
				\item 128GB of physical RAM
				\item $\approx110$GB of useable RAM
			\end{itemize}
		\item 6 large memory nodes
			\begin{itemize}
				\item Four 10 core {\intel} processors (\emph{40 cores total})
				\item Diskless -- Some RAM is used for the Operating System
				\item 1TB of physical RAM
				\item $\approx1008$GB of useable RAM
			\end{itemize}			
		\item CentOS Linux
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{{\craycs} -- Storage $\rightarrow$ {\lustre}}
	\begin{itemize}
		\item {\lustre} is a high performance parallel filesystem
		\item The {\craycs} has $\approx582$TB of storage space
		\item Shared between all compute nodes and login nodes
		\item Primarily used as scratch space for jobs (Input and Output)
		\item User do not have a usage quota (soft or hard)
		\item Certain directories are subject to a 90 day purge policy
		\item \textbf{Data is not backed up!  Users are responsible for their own data}
                \item {\lustre} utilizes RAID 6 arrays, and is fairly robust but \ldots~\\RAID is not a backup
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{{\craycs} -- Storage $\rightarrow$ ValueStorage}
	\begin{itemize}
		\item $500$TB of scale out storage
		\item Currently only available for purchase by cluster users and only accessible via the login nodes
		\item ValueStorage will eventually be available for purchase by everyone
		\item ValueStorage owners will potentially be able to mount as a network drive (CIFS) on laptop, workstations, servers
		\item Purchased in 0.5TB increments
	\end{itemize}
	\btVFill

@@

\begin{frame}
	\frametitle{{\craycs} -- Network}
	\begin{itemize}
		\item 40Gb Infiniband inter-connects (QDR)
		\begin{itemize}
			\item High speed inter-connect between~\\compute nodes, {\lustre} storage and Login nodes
                        \item low latency ($\approx1.3\mu$s)
			\item Utilizes the \emph{fat tree network topology}
			\item[] \includegraphics[width=0.20\textwidth]{images/Fat_tree_network} \\[-1ex] {\fontsize{3}{4} \selectfont Source: \url{https://en.wikipedia.org/wiki/Fat_tree} } 		
		\end{itemize}
		\item 10Gb login node internet connectivity
		\begin{itemize}
			\item Speed test from UH to CERN clocked transfer speeds up to 2$+$ Gb/s
		\end{itemize}	
	\end{itemize}
\end{frame}

@@

\section{Sustainability}
\subsection{Community Resources}
\begin{frame}
	\frametitle{Community Resources}
	\begin{itemize}
		\item The initial investment in the cluster provides resources for all faculty, staff, and students affiliated with the University of {\hawaii}
		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, lm.q, sb.q, kill.q, htc.q}
		\item For some users, the publicly available resources may not be enough~\ldots
	\end{itemize}
\end{frame}

@@

\subsubsection{Service Units}
\begin{frame}	
	\frametitle{Service Units}
	\begin{itemize}
	 \item In some cases, users will not require owning a node, but still need priority access
	 \item An alternative to purchasing a node, is to purchase service units (\emph{SU})
	 \item SUs come in two flavors:
		 \begin{itemize}
			 \item[--] Standard node units 
				\begin{itemize}
					\item 20 core hours, with access to 128GB of ram on a standard~node
				\end{itemize}
			 \item[--] Large memory node units
				\begin{itemize}
					\item 40 core hours, with access to 1TB of ram on a large~memory~node
				\end{itemize}
		 \end{itemize}
	 \item A minimum order totaling \$500 is required
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Globus}\footnotesize
\begin{block}{How do I get Globus?}
To utilize Globus, follow the following steps:
\begin{itemize}\tiny
\item Register for a Globus Online account -- \url{https://www.globusonline.org/signup}
\item Sign in to Globus Online (using your Globus Online username and password) -- \url{https://www.globusonline.org/signin}
\item Select ‘Start Transfer’ under ‘File Transfer’, or from the drop down menu in the top bar
\item You can view the list of available endpoints by clicking the button on the ‘Endpoint’ drop down box
\begin{itemize}\tiny
 \item Each of the login nodes is also an endpoint: \textbf{\emph{hawaii\#UHHPC1}} \& \textbf{\emph{hawaii\#UHHPC1}}
\end{itemize}
\item Once you select an endpoint, a login window will pop up. You can access the UHHPC endpoints by simply using your UH username and password. Enter your UH accounts username in the ‘Username’ field and UH accounts password in the ‘Password’ field and click ‘Authenticate’. 
\item You will see a listing of the contents of your home directory on the UH HPC. Double click on a directory to view its contents
\item Select a file or directory and click on the highlighted ‘arrow button’ to initiate the transfer
\end{itemize}
\end{block}
~\\
These instructions can also be found on the {\ci} website:~\\
{\footnotesize \href{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}{http://www.hawaii.edu/its/ci/hpc-resources/hpc-tutorials/globus-quick-start-guide/}}
\end{frame}

\begin{frame}
	\frametitle{Globus}
        \begin{itemize}
          \item In order to transfer data from the {\craycs} to your personal computer, a client called the Globus Connect Personal (\url{https://www.globus.org/globus-connect-personal}) needs to be installed
          \item The Globus Connect Personal, turns your personal computer into a private endpoint that is only useable with your personal Globus Online account.
          \item If you want to install Globus on your own server, Globus Connect (\url{https://www.globus.org/globus-connect-server}) is required
            \item If issues arise while trying to install Globus Connect or Globus Connect Personal,  please contact us and we will be more than happy to help
\end{itemize} 
\end{frame}

@@

\subsection{Managing User Jobs}
\begin{frame}
\frametitle{Managing User Jobs}
User jobs all come in different shapes and sizes:  
\begin{itemize}\footnotesize
	\item Require multiple nodes working in concert towards a common goal (MPI)
	\item Require a single node, in which they use multiple threads~\\work together (OpenMP, pthreads)
	\item Require a lot of cores to process a lot of data in an identical manner,~\\yet none of the inputs have dependencies on another (HTC)
\end{itemize}
\bigskip
The {\craycs} is capable of handling many different types of jobs, but with so many users in a multi-user environment, how do we impose order on this chaos?
\btVFill
\begin{center}This looks like a job for a \textbf{\emph{job scheduler}}!\end{center}
\end{frame}

\begin{frame}
	\frametitle{What is a Job Scheduler?}
	\begin{definition}{\footnotesize
		A job scheduler is a computer application for controlling unattended background program execution (commonly called batch processing)
		}
	\end{definition}
	Basic features expected of a job scheduler include:
	\begin{itemize}\footnotesize

        \item The ability for the user to interact with the schduler to define workflows and/or define job dependencies
        \item[--] Interfaces which help to define workflows and/or job dependencies

        \item The ability to automatically manage the scheduling of user jobs
        \item[--] Automatic submission of executions

        \item The functionality to allow users to monitor submitted jobs
        \item[--] Interfaces to monitor the executions

        \item The capability to control and prioritize the execution order of unrelated jobs
        \item[--] Priorities and/or queues to control the execution order of unrelated jobs
	\end{itemize}
	For the {\craycs}, we utilize the \textbf{S}imple \textbf{L}inux \textbf{U}tility \textbf{R}esource \textbf{M}anager~\\ or simply known as the \emph{SLURM scheduler}
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Job_scheduler}}
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Slurm_Workload_Manager}}
	\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/slurm.html}}
\end{frame}

@@

\subsection{SLURM Partitions}
\begin{frame}
\footnotesize
\frametitle{Partitions}
\begin{block}{What is a partition?}\footnotesize
A partition can be thought of as a group of nodes/resources divided into possibly overlapping sets.  Each partition can be considered as a job queue, each of which has an assortment of constraints such as job size limit, job time limit, users permitted to use it, etc. Priority-ordered jobs are allocated nodes within a partition until the resources (nodes, processors, memory, etc.) within that partition are exhausted.\footnotemark
\end{block}
\begin{itemize}
	\item The {\craycs} currently has six public partitions:~\\\textbf{lm.q}, \textbf{community.q}, \textbf{exclusive.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
\end{itemize}
\footnotetext[2]{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Partition Details}\tiny
\begin{tabular}{c c c c c c c}
  \toprule                                                                    
\textbf{Partition} & \textbf{Time} & \textbf{Nodes} & \textbf{Priority} & \textbf{Shared} & \textbf{Preempt\\Mode} & \textbf{Mem per CPU} \\
\midrule 
community.q & \begin{tabular}{c c}\textbf{Def}&\textbf{Max}\\00:10:00&3-00:00:00\end{tabular} & \begin{tabular}{c c}\textbf{Min}&\textbf{Max}\\1&20\end{tabular} & 10 & NO & OFF &  \begin{tabular}{c c}\textbf{Def}&\textbf{Max}\\3250&6500\end{tabular}\\
exclusive.q & \begin{tabular}{c  c}\textbf{Def}&\textbf{Max}\\00:10:00&3-00:00:00\end{tabular} &  \begin{tabular}{c  c}\textbf{Min}&\textbf{Max}\\1&20\end{tabular} & 10 & NO & OFF &  \begin{tabular}{c  c}\textbf{Def}&\textbf{Max}\\3250&6500\end{tabular}\\
\bottomrule 
\end{tabular}

%% \begin{columns}[t]
%%   \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
%% PartitionName=community.q
%%    DefaultTime=00:10:00
%%    MaxTime=3-00:00:00 
%%    MinNodes=1
%%    MaxNodes=20 
%%    Nodes=prod2-[0003-0085]
%%    Priority=10
%%    Shared=NO 
%%    PreemptMode=OFF
%%    DefMemPerCPU=3250 
%%    MaxMemPerCPU=6500
%% \end{semiverbatim}
%%   \end{column}

%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
%% PartitionName=exclusive.q
%%    DefaultTime=00:10:00 
%%    MaxTime=3-00:00:00 
%%    MinNodes=1 
%%    MaxNodes=20 
%%    Nodes=prod2-[0003-0085]
%%    Priority=10
%%    Shared=EXCLUSIVE 
%%    PreemptMode=OFF
%%    DefMemPerNode=UNLIMITED 
%%    MaxMemPerCPU=6500
%% \end{semiverbatim}
%% \end{column}

%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
%% PartitionName=lm.q
%%    DefaultTime=00:10:00
%%    MaxTime=3-00:00:00 
%%    MinNodes=1 
%%    MaxNodes=1 
%%    Nodes=lm2-[0001-0006]
%%    Priority=10
%%    Shared=NO
%%    PreemptMode=REQUEUE
%%    DefMemPerNode=UNLIMITED 
%%    MaxMemPerCPU=262100
%% \end{semiverbatim}
%% \end{column} 
%% \end{columns}

%% \begin{columns}[t]
%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
%% PartitionName=kill.q
%%    DefaultTime=00:10:00
%%    MaxTime=3-00:00:00
%%    MinNodes=1
%%    MaxNodes=12
%%    Nodes=prod2-[0085-0176]
%%    Priority=10
%%    Shared=NO
%%    PreemptMode=REQUEUE
%%    DefMemPerCPU=3250
%%    MaxMemPerCPU=6500

%% \end{semiverbatim}
%% \end{column}

%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
%% PartitionName=htc.q
%%    DefaultTime=00:10:00
%%    MaxTime=3-00:00:00 
%%    MinNodes=1 
%%    MaxNodes=1 
%%    Nodes=prod2-[0001-0176]
%%    Priority=1
%%    Shared=NO
%%    PreemptMode=REQUEUE
%%    DefMemPerCPU=3250 
%%    MaxMemPerCPU=6500
%% \end{semiverbatim}
%% \end{column}

%% \begin{column}{.3\textwidth}\begin{semiverbatim}\tiny
%% PartitionName=sb.q
%%    DefaultTime=00:05:00
%%    MaxTime=01:00:00
%%    MinNodes=1
%%    MaxNodes=2
%%    Nodes=prod2-[0177-0178]
%%    Priority=10 
%%    Shared=NO 
%%    PreemptMode=REQUEUE
%%    DefMemPerCPU=3250 
%%    MaxMemPerCPU=6500
%% \end{semiverbatim}
%% \end{column} 
%% \end{columns}

More details on partition constraints can be viewed by visiting the \href{http://www.hawaii.edu/its/ci/hpc-resources/slurm-partitions/}{{\ci} website} or by using the following command
\begin{itemize}\footnotesize
\item[] \texttt{$[$login \ctilde$]$\$ scontrol show partition $<$partition name$>$}
\item[] Example: \texttt{$[$login \ctilde$]$\$ scontrol show partition community.q}
\end{itemize}
